### 2022-12-21 Manuscript/Draft1.Rmd
update manuscript

---
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
  bibliography: IBMWG.bib
  word_document: default
  html_document: default
  pdf_document:
    toc: no
fontsize: 12pt
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
- \usepackage[belowskip=0pt,aboveskip=0pt]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(kableExtra)) {  
  install.packages("kableExtra")
  require(kableExtra)}

@@

```{r}
knitr::kable(YT_results, caption = "Yellowtail error results")

@@

Fish stocks in this highly productive and economically important region are managed by the National Oceanic and Atmospheric Administration’s (NOAA) Northeast Fisheries Science Center (NEFSC) in Woods Hole, Massachusetts. Federal biologists assess the health and abundance of each commercial fish stock using fishery-independent bottom trawl survey data that has been collected by NOAA throughout the region since 1963 (cite survey paper). The survey uses a stratified random design where bottom trawl sampling takes place in predefined strata along the eastern continental shelf **(include image from johnson_sosebee?)**. The survey has created a rich time series data set with many uses including species-specific habitat identification, analysis of how environmental conditions influence species abundance, and estimating yearly species abundance trends to help inform stock assessments and ultimately quota limits **just listed a few uses of survey change/others?**.

@@

We obtained a shapefile for the 15 strata that comprise George's Bank to use as our modeling environment. We discritized the region into a raster with 88 rows and 144 columns. Haddock inhabit all 15 strata in the domain Cod inhabit 13 strata, and yellowtail exist in 9 strata. **include figure?**

@@

The time step for our models is one week. MFS uses a modified two-stage Deriso-Schnute delay difference equation that models the biomass of in each cell in our study area [@dolder2020highly]. Individual terms in the formulation account for growth of mature adults, natural and fishing mortality, and the addition of new recruits. We chose to represent recruitment in the model using a Beverton-Holt formulation. Recruitment is a function of the adult biomass that existed in the previous year and is added to the population incrementally throughout each species' predefined spawning period.

--Many parameter inputs were obtained from literature

Each species is assumed to have normally distributed temperature preferences ($~N(\mu,\sigma)$). We assume Yellowtail's preferences are $N(8.75,4.25)$, while Haddock and Cod have preferences $N(9,4)$. We chose these values by combining information in the literature with temperatures recorded in the bottom trawl survey. A full list of parameters used in our model can be seen below in Table ???.

**go through each process in model and discuss edits/removals?**

@@

The package was designed to generate theoretical habitat preferences using Gaussian Random Fields that combine with hypothetical temperature gradients to simulate an entire fishery. Since we are modeling real species on the northeast continental shelf w

We obtain estimated temperature data for the region for 2012 from FVCOM **cite**.

@@

We consider three population parameter scenarios for each of our three species- a scenario where parameters result in each population increasing over time, one where the populations are relatively constant over time, and a scenario where the parameter combination results in each population decreasing over time. Each of these three scenarios is paired with a temperature gradient that repeats as well as one that increasing roughly 5 degrees calcius over the duration of the 20 year simulation.

@@

After each simulation is complete, we mimic the bottom trawl survey by conducting stratified random sampling in each inhabited strata twice each year. The number of the samples taken in each strata reflect true values and we sample from the same weeks that the Spring and Fall surveys take place. Most strata contain enough cells to sample a unique location in each survey. For smaller strata we are forced to repeat locations. We then use the samples in contemporary population indexing methods to track population trends. Since we know the true population values in our simulations we can compare methods by comparing the error calculated from each estimate.

@@

We compare the stratified mean estimate of abundance to estimates obtained from the Vector-Autoregressive Spatio-temporal (VAST) model. The stratified mean is a typical survey-based approach that scales individual samples to the  strata-level by considering the area of each strata, before scaling to the region-level based on the relative size of each strata. VAST is a spatio-temporal statistical framework that models both abundance (biomass) and probability of occurance (presence/absence). If desired, VAST also allows users to include covariate data to better inform the model.

@@

\pagebreak
Table 1. Parameters used in all population models.
```{r paramsALL, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|		|	Description | Unit |Yellowtail | Cod | Haddock | Source
|:---	|:------- |:-- |:--- |:-- |:-- |:--
| $\\rho$ | Ford's growth coefficient | wk$^{-1}$  | 4.48 | 4.43 | 4.49 | |
| $M$ | Natural Mortality| wk$^{-1}$  | 0.2064 | 0.2728 | 0.3340 | |
| $W_R$ |Weight of fully recruited fish| kg | 0.39 | 2.95 | 1.12 | |
| $W_{R-1}$ | Weight of pre-recruit fish| kg | 0.13 | 0.39 | 0.19 | |
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

Table 2. Parameters used in models with relatively constant populations.
```{r paramsCON, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|		|	Description | Yellowtail | Cod | Haddock 
|:---	|:------- |:--- |:-- |:-- |:--
| $M+F$ | Adjusted Mortality (Natural + Fishing)| wk$^{-1}$  | 0.764 | 0.83 | 0.309 |

"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

Table 3. Parameters used in models with increasing populations.
```{r paramsINC, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|		|	Description | Yellowtail | Cod | Haddock | Source
|:---	|:------- |:--- |:-- |:-- |:--
| $M+F$ | Adjusted Mortality (Natural + Fishing)| wk$^{-1}$  | 0.564 | 0.372 |  0.134 |

"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

Table 4. Parameters used in models with decreasing populations.
```{r paramsDEC, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|		|	Description | Yellowtail | Cod | Haddock | Source
|:---	|:------- |:--- |:-- |:-- |:--
| $M+F$ | Adjusted Mortality (Natural + Fishing)| wk$^{-1}$  | 0.764 | 0.623 | 0.334 |

"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

>>>>

---
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
  bibliography: IBMWG.bib

  word_document: default
  html_document: default
  pdf_document:
    toc: no
fontsize: 12pt
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
- \usepackage[belowskip=0pt,aboveskip=0pt]{caption}
 \usepackage{array}
 \usepackage{caption}
 \usepackage{graphicx}
 \usepackage{siunitx}
 \usepackage{colortbl}
 \usepackage{multirow}
 \usepackage{hhline}
 \usepackage{calc}
 \usepackage{tabularx}
 \usepackage{tabulary}
 \usepackage{threeparttable}
 \usepackage{wrapfig}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
if(!require(kableExtra)) {  
  install.packages("kableExtra")
  require(kableExtra)}

@@

```{r}

knitr::kable(YT_results, caption = "Yellowtail error results")

@@

Fish stocks in this highly productive and economically important region are managed by the National Oceanic and Atmospheric Administration’s (NOAA) Northeast Fisheries Science Center (NEFSC) in Woods Hole, Massachusetts. Federal biologists assess the health and abundance of each commercial fish stock using fishery-independent bottom trawl survey data that has been collected by NOAA throughout the region since 1963 (cite survey paper). The survey uses a stratified random design where bottom trawl sampling takes place in predefined strata along the eastern continental shelf. The survey has created a rich time series data set with many uses including species-specific habitat identification, analysis of how environmental conditions influence species abundance, and estimating yearly species abundance trends to help inform stock assessments and ultimately quota limits **just listed a few uses of survey change/others?**.

@@

We obtained a shapefile for the 15 strata that comprise George's Bank to use as our modeling environment. We discritized the region into a raster with 88 rows and 144 columns. Haddock inhabit all 15 strata in the domain Cod inhabit 13 strata, and yellowtail exist in 9 strata. Figure \@ref(fig:strata-plot) shows the regions used in our models.

```{r strata-plot, echo=FALSE,out.width="95%",fig.show='hold', fig.align='center', fig.cap="Strata inhabited by each species in our population models."}
knitr::include_graphics(c("Images/Strata.png"))
```

@@

The time step for our models is one week. MFS uses a modified two-stage Deriso-Schnute delay difference equation that models the biomass of in each cell in our study area [@dolder2020highly]. Individual terms in the formulation account for growth of mature adults, natural and fishing mortality, and the addition of new recruits. We chose to represent recruitment in the model using a Beverton-Holt formulation. Recruitment is a function of the adult biomass that existed in the previous year and is added to the population incrementally throughout each species' predefined spawning period. Parameter inputs were either obtained from the literature or chosen to produce desired model dynamics. A full list of parameters used in our model can be seen below in Tables \@ref(paramsALL) and \@ref(tab:paramsSCENARIOS). **missing spawning weeks, recruitment weeks, lambda**

@@

The package was designed to generate theoretical habitat preferences using Gaussian Random Fields that combine with hypothetical temperature gradients to drive the probability of movement from cell $I$ to cell $J$ using the formulation

\begin{align*}
Pr(C_{wk+1}=J|C_{wk}=I) = \frac{e^{-\lambda \cdot d_{I,J}}\cdot(Hab^2_{J,p} \cdot Tol_{J,p,wk})}{\sum^C_{c=1}e^{-\lambda \cdot d} \cdot (Hab^2_{c,p} \cdot Tol_{c,p,wk})},
\end{align*}

 where

   $e^{-\lambda \cdot d_{I,J}}$ accounts for distance between cells $I$ and $J$,

  $Hab^2_{J,p}$ is the static habitat value for species $p$ in cell $J$, and

   $Tol_{c,p,wk}$ is the value from normally distributed temperature tolerance for species $p$ in cell $c$ in week $wk$.

Since we are modeling real species on the northeast continental shelf, we formulate the habitat and temperature components as follows.

Each species is assumed to have normally distributed temperature preferences ($~N(\mu,\sigma)$). We assume Yellowtail's preferences are $N(8.75,4.25)$, while Haddock and Cod have preferences $N(9,4)$. We chose these values by combining information in the literature with temperatures recorded in the bottom trawl survey. We obtain estimated temperature data for the region for 2012 from FVCOM **cite**. We chose 2012 because the data displayed an average temperature pattern that consistently oscillated between maximum and minimum temperature values. We transform the temperature data to create an oscillating  pattern that increases 5 degrees Celsius over the duration of the simulation. **show image of average temperature oscilations?**

Species-specific habitat preferences were derived by relating bottom trawl data to covariate predictors to create a niche model for each species.

@@

We consider three population parameter scenarios for each of our three species- a scenario where parameters result in each population increasing over time, one where the populations are relatively constant over time, and a scenario where the parameter combination results in each population decreasing over time. Each of these three scenarios is paired with a temperature gradient that repeats as well as one that increasing roughly 5 degrees Celsius over the duration of the 20 year simulation.

@@

After each simulation is complete, we mimic the bottom trawl survey by conducting stratified random sampling in each inhabited strata twice each year. We sample in the same weeks that the Spring and Fall surveys take place and the number of the samples taken in each strata reflect true values. Most strata contain enough cells to sample a unique location in each survey over the duration of the simulation. For smaller strata we must repeat some sample locations. We then use the biomass collected from our samples in contemporary population indexing methods to estimate population trends. Knowing the true population values in our simulations allows us to compare the error calculated from each estimation method.

@@

We compare the stratified mean estimate of abundance to estimates obtained from the Vector-Autoregressive Spatio-temporal (VAST) model. The stratified mean is a typical survey-based approach that scales individual samples to the  strata-level by considering the area of each strata, before scaling to the region-level based on the relative size of each strata. VAST is a spatio-temporal statistical framework that models both abundance (biomass) and probability of occurrence (presence/absence). If desired, VAST also allows users to include covariate data to better inform the model.

@@

\pagebreak
Table 1. Parameters used in all population models.
```{r paramsALL, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|		|	Description | Unit |Yellowtail | Cod | Haddock | Source
|:---	|:------- |:-- |:--- |:-- |:-- |:--
| $\\rho$ | Ford's growth coefficient | wk$^{-1}$  | 4.48 | 4.43 | 4.49 | |
| $M$ | Natural Mortality| wk$^{-1}$  | 0.2064 | 0.2728 | 0.3340 | |
| $W_R$ |Weight of fully recruited fish| kg | 0.39 | 2.95 | 1.12 | |
| $W_{R-1}$ | Weight of pre-recruit fish| kg | 0.13 | 0.39 | 0.19 | |
| $\\sigma^2$ | Variance in recruited fish | kg$^2$ | 0.55 | 0.55 | 0.55 ||
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

<!-- Table 2. Parameters used in models with relatively constant populations. -->
<!-- ```{r paramsCON, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->
<!-- tabl <- " -->
<!-- |		|	Description | Yellowtail | Cod | Haddock  -->
<!-- |:---	|:------- |:--- |:-- |:-- |:-- -->
<!-- | $M+F$ | Adjusted Mortality (Natural + Fishing)| wk$^{-1}$  | 0.764 | 0.83 | 0.309 |  -->

<!-- " -->
<!-- cat(tabl) # output the table in a format good for HTML/PDF/docx conversion -->
<!-- ``` -->

<!-- Table 3. Parameters used in models with increasing populations. -->
<!-- ```{r paramsINC, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->
<!-- tabl <- " -->
<!-- |		|	Description | Yellowtail | Cod | Haddock | Source -->
<!-- |:---	|:------- |:--- |:-- |:-- |:-- -->
<!-- | $M+F$ | Adjusted Mortality (Natural + Fishing)| wk$^{-1}$  | 0.564 | 0.372 |  0.134 |  -->

<!-- " -->
<!-- cat(tabl) # output the table in a format good for HTML/PDF/docx conversion -->
<!-- ``` -->

<!-- Table 4. Parameters used in models with decreasing populations. -->
<!-- ```{r paramsDEC, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->
<!-- tabl <- " -->
<!-- |		|	Description | Yellowtail | Cod | Haddock | Source -->
<!-- |:---	|:------- |:--- |:-- |:-- |:-- -->
<!-- | $M+F$ | Adjusted Mortality (Natural + Fishing)| wk$^{-1}$  | 0.764 | 0.623 | 0.334 |  -->

<!-- " -->
<!-- cat(tabl) # output the table in a format good for HTML/PDF/docx conversion -->
<!-- ``` -->

<!-- DOESNT WORK!!
<!-- Table 2. Parameters used in population models for each scenario. -->
<!-- ```{r paramsSCENARIOS, echo=FALSE, message=TRUE, warnings=TRUE, results='asis'} -->
<!-- library(huxtable) -->

<!-- ht <- tribble_hux( -->
<!--   ~ "Section/topic", ~ "Item no", ~ "Checklist item", ~ "Reported on Page No", -->
<!--   "Title"          , ""         , ""                , "", -->
<!--   "Title"          , "1"        , "Identify..."     , "",  -->
<!--   "Abstract"       , ""         , ""                , "", -->
<!--   "Structured summary", "2"     , "Provide..."      , "" -->
<!--   # et cetera... -->
<!-- )  -->

<!-- # using the pipe from 4.1.0... -->

<!-- ht |>  -->
<!--   set_header_rows(c(2, 4), TRUE)                |> -->
<!--   merge_across(c(2, 4), everywhere)             |> -->
<!--   style_header_rows(bold = TRUE)                |> -->
<!--   set_all_borders(brdr(0.4, "solid", "grey70")) |> -->
<!--   set_background_color("grey97")                |> -->
<!--   set_background_color(1, 1:3, "grey90")        |> -->
<!--   set_col_width(c(0.2, 0.05, 0.55, 0.2))        |> -->
<!--   set_font("cmss")                               -->
<!-- ``` -->

<!-- ```{r,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->
<!-- library("knitr") -->
<!-- library("kableExtra") -->
<!-- library("dplyr") -->
<!-- library("readr") -->
<!-- library("tidyr") -->
<!-- library("forcats") -->

<!-- #orig_ui_deaths = read_csv("https://www.opendata.nhs.scot/dataset/b0135993-3d8a-4f3b-afcf-e01f4d52137c/resource/89807e07-fc5f-4b5e-a077-e4cf59491139/download/ui_deaths_2020.csv") -->

<!-- summary_tab <- orig_ui_deaths %>% -->
<!--   # Apply filters in same was as in plots so looking at one year, whole of Scotland, and exlucing "All" entries -->
<!--   filter(Year == "2018", HBR == "S92000003", AgeGroup != "All" & Sex != "All", InjuryType != "Accidental exposure" & InjuryType != "All") %>% -->
<!-- # Group the table according to injury type, age, and sex -->
<!--   group_by(InjuryType, AgeGroup, Sex) %>% -->
<!-- # Create a summary of total number of deaths for neater appearance to the table and demonstration of the figures -->
<!--   summarise(total_deaths = sum(NumberofDeaths)) %>% -->
<!-- # Change the orientation of the table so that age groups become the variable headings -->
<!--   pivot_wider(names_from = AgeGroup, values_from = total_deaths) %>% -->
<!--   mutate(Sex = factor(Sex)) %>% -->
<!--   arrange(Sex) %>% -->
<!--   select(Sex, everything()) -->

<!-- # Create table to present the data] -->
<!-- Fig_2 <- kable(summary_tab[, -1],  -->
<!--   caption = "Deaths from unintentional injuries in Scotland", -->
<!--     format = "latex", booktabs = TRUE) %>% -->
<!--   kable_styling(font_size = 10) %>% -->
<!--   pack_rows(tab_kable, colnum = 1, -->
<!--     index = table(fct_inorder(summary_tab$Sex), useNA = "no")) -->
<!-- ``` -->
<!-- ```{r, results='asis'} -->
<!-- Fig_2 -->
<!-- ``` -->

```{r paramsSCENARIOS,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

param_table = read_csv("Parameters.csv")

summary_tab <- param_table %>% 
  dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3))))%>%

 mutate_if(grepl('alpha',.), ~replace(., grepl('alpha', .), paste("\U03B1"))) %>% #should be replacing alpha with the symbol
  mutate_if(grepl('beta',.), ~replace(., grepl('beta', .), paste("\U03B2"))) %>% #should be replacing beta with the symbol
    mutate_if(grepl('wk',.), ~replace(., grepl('wk', .), "1/wk")) %>% #should be replacing beta with the symbol

# Group the table according to Parameter, Scenario
  group_by(Parameter, Scenario)%>%
# Change the orientation of the table so that Species become the variable headings
  pivot_wider(names_from = Species,  values_from = Value) %>%
  mutate(Scenario = factor(Scenario)) %>%
  arrange(Scenario) %>%
  select(Scenario, everything())

# Create table to present the data]
Fig_3 <- kable(summary_tab[, -1], 
  caption = "Parameters used in population models for each scenario.",
    format = "latex", booktabs = TRUE) %>%
  kable_styling(font_size = 10) %>%
  pack_rows(tab_kable, colnum = 1,
    index = table(fct_inorder(summary_tab$Scenario), useNA = "no"))
```
```{r, results='asis'}
Fig_3
```
